---
banner: img/banners/symountains.jpg
title: "Gibraltar Reservoir Capacity"
author: "Jessica Jagdeo, Lydia Bleifuss, Camila Bobroff"
date: "10/30/2019"
output: html_document
categories: ["R"]
tags: ["R Markdown", "plot", "hugo", "water"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

Construction of Gibraltar Reservoir finished in 1920. The capacity of the reservoir was supposed to be slightly over 20,000 acre-feet, but sedimentation had already reduced the water storage capacity of the reservoir to about 15,000 acre-feet. Since 1920, Gibraltar has faced issues from sedimentation. In 1949, the City of Santa Barbara had to raise Gibraltar Dam due to the amount of sediment accumulating within its confines. Raising the dam maintained the water storage capacity of about 15,000 acre-feet. This project visualizes annual water storage and total water storage capacity of Gibraltar Reservoir. 

#### 1. Load in necessary packages
```{r}

library(tidyverse)
library(readxl)
library(janitor)
library(lubridate)
library(gghighlight)
library(kableExtra)
library(ggplot2)
library(ggthemes)
library(here)

```

#### 2. Load in the data and clean
```{r}

gibraltar_storage <- read_excel("RAW_Gibraltar_Storage_Elevation_2001_2019.xls") %>% 
  rename("date" = "36892", "capacity" = "3182", "total_capacity" = "7500") %>%
  mutate(capacity_m3 = (capacity*1233.48))

```

## Water storage in Gibraltar Reservoir

```{r}

gibraltar_cap_plot <- ggplot(gibraltar_storage, aes(x = date, y = capacity)) +
  geom_area(color = "darkblue", fill = "darkblue") + # creates area rather than bars
  theme_classic() +
  labs(x = "Year", 
       y = "Water (acre-feet)"
       ) +
  ggtitle("Gibraltar Water Stored (2001 - 2018)\n") +
  theme(plot.title = element_text(hjust=0.5)) +
  theme(plot.title = element_text(size=12)) +
  scale_y_continuous(expand = c(0,0),
                     limits = c(0, 15000),
                     labels = scales::comma, 
                     breaks = c(5000, 10000, 15000)) + 
  geom_hline(yintercept = 7500, linetype = "dashed", size = .8, color = "black")  # Total possible water capacity was 7500 acre-feet in 2001 (when the data in this frame begins) 
  #theme_black()
  
gibraltar_cap_plot

```

**Figure 1.** Historical Water Storage for Gibraltar 2001 - 2018. Source: Santa Barbara County, Public Works Department.

```{r}


gibraltar_total_bathymetric <- read_excel("RAW_Table of Bathymetric Data at 1400 Elevation.xlsx") %>% 
  clean_names() %>% 
  slice(1:23) %>%  # Kept years and capacity data & excluded irrelevant data 
  mutate(sed_inflow = case_when(
    year < 1945 ~ 14500 - volume_in_af,
    year > 1945 ~ 14800 - volume_in_af
  )) %>% # Used case_when to apply the different original capacities. For 1920 and 1944, the total original capacity is 14,500. For years after 1944, the total original capacity is 14,800 because the dam was raised (JJ)
  mutate(sed_inflow_m3 = (sed_inflow*1233.48)) %>%  # Converting from acre-feet to cubic meters
  filter(volume_in_af != 8939.00) # Removing the May 1979 survey because there were 2 bathymetric surveys conducted in 1979. When we plot sed in the following graph, the volumes for each survey are added together, making 1979 unusually large (JJ)

gibraltar_total_bathymetric

gibraltar_total_bathymetric$year = as.numeric(gibraltar_total_bathymetric$year)

#note that for 1920 and 1944, the total original capacity should be 14500 and not 14800 
```

## Sedimentation in Gibraltar Reservoir

```{r}

cumulative_bathymetric_inflow <- gibraltar_total_bathymetric %>% 
  mutate(cumulative_sed_inflow_m3 = case_when(
    year <= 1944 ~ sed_inflow_m3,
    year > 1944 ~ sed_inflow_m3 + 8511012))

cumulative_bathymetric_inflow <- cumulative_bathymetric_inflow %>% 
  mutate(cumulative_sed_inflow_af = cumulative_sed_inflow_m3/1233)

gibraltar_vol_graph <- ggplot(cumulative_bathymetric_inflow) +
  # geom_bar(stat="identity", width = 1)+
  geom_area(aes(x = year, y = cumulative_sed_inflow_af), color = "burlywood4", fill = "burlywood4", alpha = 0.6) +
  labs(x = "Year", 
       y = "Volume (acre-feet)") +
  ggtitle("Gibraltar Reservoir Capacity Cross Section (1920 - 2019)")+
  theme(plot.title = element_text(size=12))+
  theme(plot.title = element_text(hjust=0.5)) +
   scale_y_continuous(expand = c(0,0),
                      limits = c(0,25000)) +
  geom_segment(aes(x = 1920, y = 15322.502189, xend = 1949, yend = 15322.502189)) + # original design capacity of Gibraltar Reservoir was 15,374 acre feet
  geom_segment(aes(x = 1949, y = 22537.860363, xend = 2019, yend = 22537.860363)) + # After its 1948 expansion the gross capacity was increased to 22,500 acre feet
  geom_segment(aes(x = 1949, y = 15322.502189, xend = 1949, yend = 22537.860363)) +
  scale_x_continuous(breaks = c(1920,1940,1960,1980,2000,2020),
                     expand = c(0,0)) +
  theme(panel.background = element_rect(fill = "lightblue",
                                colour = "lightblue"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) 
  

gibraltar_vol_graph

```

**Figure 2.** Reservoir Capacity of Gibraltar 1920 - 2019. The tan area indicates sediment accumulation within the reservoir, whereas the blue area indicates available water capacity. The solid black line represents the total capacity of the reservoir, including sedimentation. In 1949, the dam was raised, increasing the capacity of Gibraltar Reservoir. Source: City of Santa Barbara, Public Works Department.