---
banner: img/banners/beach.jpg
title: "Examining Runoff Ratio in Santa Cruz Creek Sub-basin"
author: "Jessica Jagdeo & Lydia Bleifuss"
date: "1/13/2020"
output: html_document
categories: ["R"]
tags: ["R Markdown", "plot", "hugo", "runoff"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

To understand the hydrological processes specific to a particular watershed, it is important to understand how much runoff is generated within the area. The runoff ratio for a particular watershed can be calculated if there are good records of precipitation and discharge within the watershed. In the Santa Ynez Watershed, Santa Barbara County maintains a record of precipitation at Lake Cachuma and the US Geological Survey records discharge at the Santa Cruz sub-basin, which is near Lake Cachuma. Using these data sets, the runoff ratio for the Santa Cruz sub-basin was calculated from 1983 to 2018 for the rainy season. 

#### 1. Load necessary packages.
```{r}

library(tidyverse)
library(readxl)
library(lubridate)
library(janitor)
library(ggthemes)
library(zoo)
library(magrittr)
library(gghighlight)
library(tsibble)
library(paletteer)
```

#### 2. Load in the data and clean

Santa Cruz Gauge Observed Flow (US Geological Survey)
```{r}

# Read in SC observed flow data

sc_usgs_flow <- read_csv("sc_creek_clean.csv")

sc_usgs_flow$dis_cfs <- as.numeric(as.character(sc_usgs_flow$dis_cfs))

# Convert date from character to date

sc_usgs_flow$date <- as.Date(sc_usgs_flow$date, "%m/%d/%Y")

# Create column for daily inflow in cumecs (cfs to cubic meters per second (cumecs)) 

sc_flow_cumecs <- sc_usgs_flow %>% 
  mutate(dis_cumecs = (dis_cfs*0.03)) %>% 
  select(date, dis_cumecs)

```

#### 3. Surface runoff calculation

Divide aggregated monthly discharge by area (km) to obtain to meters of runoff. 
Compare this runoff to rainfall (mm) per month for subbasin 4.
Try it for early and late months and high rainfall years and low rainfall years.

```{r}

options(scipen=999) #This code turns off scientific notation display in dataframes

runoff_calc <- sc_flow_cumecs %>% 
  separate(date, sep = "-", into = c("year", "month", "day")) %>% 
  group_by(year, month) %>% 
  mutate(runoff_ms = dis_cumecs/201300000) %>% #201.3 km^2 is the area of the subbasin 4. Converted it to meters to obtain runoff in meters per second.
  mutate(runoff_md = runoff_ms*86400) %>%  #Converting runoff from meters per second to meters per day by multiplying by 86400 seconds in a day
  mutate(runoff_mmd = runoff_md*1000) %>%  #Converting runoff from meters per day to millimeters per day for easy precip comparison
  summarize(
    monthly_runoff_mmd = sum(runoff_mmd)
  ) 
  

```

#### 4. Subbasin 4 Precipitation Data (Santa Barbara County Cachuma Station)
```{r}

cachuma_precip <- read_xls("RAW_332dailys_SBcounty.xls") %>%
  tail(-8) %>% 
  row_to_names(row_number = 1) %>% 
  clean_names() %>%
  filter(year >= "1983" & year <= "2018") 
  
cachuma_precip$daily_rain<- as.numeric(cachuma_precip$daily_rain)

cachuma_precip$date <- as.Date(with(cachuma_precip, paste(year, month, day, sep="-")), "%Y-%m-%d")

cachuma_precip_calc <- cachuma_precip %>% 
  select(date, daily_rain) %>% 
  mutate(daily_rain_mm = daily_rain * 25.4) %>% 
  separate(date, sep = "-", into = c("year", "month", "day")) %>%
  group_by(year, month) %>%
  summarize(
    monthly_precip_mm = sum(daily_rain_mm)
  )
         
```

#### 5. Combine monthly runoff and precip into one dataframe (insert 0s for months without rain)
```{r}

monthly_precip_runoff <- merge(cachuma_precip_calc, runoff_calc) %>% 
  mutate(runoff_precip_ratio = monthly_runoff_mmd/monthly_precip_mm)

monthly_precip_runoff$date <- as.yearmon(paste(monthly_precip_runoff$year, monthly_precip_runoff$month), "%Y %m")

monthly_precip_runoff$date <- as.Date(monthly_precip_runoff$date)


```

#### 6. Graph the runoff ratio for the Santa Cruz sub-basin during rainy months from 1983 to 2018
```{r}

selected_seasons_plot <- monthly_precip_runoff %>% 
  filter(month == "01"|
           month == "02"|
           month == "11"|
           month == "12") %>% 
  ggplot(aes(x = date, y = runoff_precip_ratio)) +
  geom_line(aes(color = month)) +
  theme_classic() +
  scale_color_paletteer_d("rcartocolor::Bold") +
  scale_color_discrete("Month", labels = c("January", "February", "November", "December")) +
  scale_y_continuous(limits = c(0,1),
                     expand = c(0,0)) +
  scale_x_date(date_breaks = "5 year", date_labels = "%Y",
               expand = c(0,0)) +
  labs(x = "Year",
       y = "Runoff Ratio") +
  ggtitle("Santa Ynez Watershed Rainy Months Runoff Ratio (1983 - 2018)") +
  theme(plot.title = element_text(hjust = 0.5, size = 12))

selected_seasons_plot

```
