---
banner: img/banners/moremesa.jpg
title: "Assessment of Santa Ynez River Water Quality"
author: "Jessica Jagdeo"
date: "11/24/2019"
output: html_document
categories: ["R"]
tags: ["R Markdown", "plot", "hugo", "waterquality"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

The quality of water within a watershed affects people who rely on surficial water for drinking, as well as the local wildlife. The Santa Ynez River flows into the Gibraltar Reservoir and Lake Cachuma and water is diverted from these reservoirs to nearby cities. The Santa Ynez Watershed supplies water to residents within the City of Santa Barbara, Goleta, Montecito, and other towns. Since 1998, water quality measurements have been taken along the Santa Ynez River. This project visualizes the turbidity and amount of total dissolved solids and total suspended solids recorded by water samples. 

#### 1. Load necessary packages.

```{r}

library(tidyverse)
library(readxl)
library(lubridate)
library(janitor)
library(ggthemes)
library(zoo)
library(ggplot2)
library(scales)

```


#### 2. Load in the data and clean

Santa Ynez Water Quality Historical Data
Data Source: 
Turbidity units: NTU
TSS and TDS units: mg/L
```{r}

santa_ynez_water_quality_raw <- read_csv("RAW_santa_ynez_water_quality.csv")

santa_ynez_water_quality <- santa_ynez_water_quality_raw %>% 
  select(OrganizationFormalName, ActivityStartDate, MonitoringLocationIdentifier, CharacteristicName, ResultSampleFractionText, ResultMeasureValue, "ResultMeasure/MeasureUnitCode") %>%  # Selecting only relevant columns 
  filter(CharacteristicName == "Total dissolved solids"|
           CharacteristicName == "Turbidity" |
           CharacteristicName == "Total suspended solids") %>% 
  separate(ActivityStartDate, sep = "-", into = c("year", "month", "day")) %>% 
   group_by(year, month, CharacteristicName) %>% 
  summarize(value = sum(ResultMeasureValue)) 

santa_ynez_water_quality$date <- as.yearmon(paste(santa_ynez_water_quality$year, santa_ynez_water_quality$month), "%Y %m")

```


#### 3. Graph turbidity, total suspended solids, and total dissolved solids samples
```{r}

santa_ynez_historical_quality <- santa_ynez_water_quality %>% 
  ggplot(aes(x = date, y = value)) +
  geom_point(color = "darkgreen") +
  facet_wrap(~CharacteristicName, scales = "free") +
  theme_classic() +
  xlab("\nYear") +
  ylab("Sampling Result\n") +
  ggtitle("Santa Ynez Watershed Turbidity, Total Suspended Solids, & Total Dissolved Solids\nMonthly 1998 - 2018") +
  theme(plot.title = element_text(hjust=0.5)) +
  theme(plot.title = element_text(size=12))
  
  santa_ynez_historical_quality

```

**Figure 1.** Monthly sampling results of turbidity, total suspended solids, and total dissolved solids for the Santa Ynez Watershed from 1998 to 2018. Turbidity units in NTU and TSS and TDS units in mg/l. Data source: California State Water Resources Control Board, Environmental Protection Agency, Santa Ynez Chumash Environmental Office. 