---
banner: img/banners/banner-1.jpg
title: "Examining Runoff Ratio in Santa Cruz Creek Sub-basin"
author: "Jessica Jagdeo"
date: "1/13/2020"
output: html_document
categories: ["R"]
tags: ["R Markdown", "plot", "hugo", "runoff"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```


```{r}

library(tidyverse)
library(readxl)
library(lubridate)
library(janitor)
library(ggthemes)
library(zoo)
library(magrittr)
library(gghighlight)
library(tsibble)
library(paletteer)
```

*Santa Cruz Gauge Observed Flow*
```{r}

# Read in SC observed flow data

sc_usgs_flow <- read_csv("sc_creek_clean.csv")

sc_usgs_flow$dis_cfs <- as.numeric(as.character(sc_usgs_flow$dis_cfs))

# Convert date from character to date

sc_usgs_flow$date <- as.Date(sc_usgs_flow$date, "%m/%d/%Y")

# Create column for daily inflow in cumecs (cfs to cubic meters per second (cumecs)) 

sc_flow_cumecs <- sc_usgs_flow %>% 
  mutate(dis_cumecs = (dis_cfs*0.03)) %>% 
  select(date, dis_cumecs)

```

*Surface runoff calculation*

Divide aggregated monthly discharge by area (km) to obtain to meters of runoff. 
Compare this runoff to rainfall (mm) per month for subbasin 4.
Try it for early and late months and high rainfall years and low rainfall years.

```{r}

options(scipen=999) #This code turns off scientific notation display in dataframes

runoff_calc <- sc_flow_cumecs %>% 
  separate(date, sep = "-", into = c("year", "month", "day")) %>% 
  group_by(year, month) %>% 
  mutate(runoff_ms = dis_cumecs/201300000) %>% #201.3 km^2 is the area of the subbasin 4. Converted it to meters to obtain runoff in meters per second.
  mutate(runoff_md = runoff_ms*86400) %>%  #Converting runoff from meters per second to meters per day by multiplying by 86400 seconds in a day
  mutate(runoff_mmd = runoff_md*1000) %>%  #Converting runoff from meters per day to millimeters per day for easy precip comparison
  summarize(
    monthly_runoff_mmd = sum(runoff_mmd)
  ) 
  

```

*Subbasin 4 Precipitation Data (Cachuma Station)*
```{r}

cachuma_precip <- read_xls("RAW_332dailys_SBcounty.xls") %>%
  tail(-8) %>% 
  row_to_names(row_number = 1) %>% 
  clean_names() %>%
  filter(year >= "1983" & year <= "2018") 
  
cachuma_precip$daily_rain<- as.numeric(cachuma_precip$daily_rain)

cachuma_precip$date <- as.Date(with(cachuma_precip, paste(year, month, day, sep="-")), "%Y-%m-%d")

cachuma_precip_calc <- cachuma_precip %>% 
  select(date, daily_rain) %>% 
  mutate(daily_rain_mm = daily_rain * 25.4) %>% 
  separate(date, sep = "-", into = c("year", "month", "day")) %>%
  group_by(year, month) %>%
  summarize(
    monthly_precip_mm = sum(daily_rain_mm)
  )
         
```

Combine monthly runoff and precip into one dataframe (insert 0s for months without rain)
```{r}

monthly_precip_runoff <- merge(cachuma_precip_calc, runoff_calc) %>% 
  mutate(runoff_precip_ratio = monthly_runoff_mmd/monthly_precip_mm)

monthly_precip_runoff$date <- as.yearmon(paste(monthly_precip_runoff$year, monthly_precip_runoff$month), "%Y %m")

monthly_precip_runoff$date <- as.Date(monthly_precip_runoff$date)


```

Graph it
```{r}

monthly_ratio <- ggplot(monthly_precip_runoff, aes(x = date, y = runoff_precip_ratio)) +
  geom_point() +
  geom_line() +
  theme_minimal() +
  labs(x = "Year",
       y = "Runoff/Precipitation Ratio") +
  ggtitle("Runoff & Precipitation Ratio for Santa Ynez Watershed") +
  scale_y_continuous(limits = c(0,1)) +
  scale_x_date(date_breaks = "12 month", date_labels = "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))

monthly_ratio

```

Seasonality graph for all months:
```{r}

seasonal_runoff_plot <- ggplot(monthly_precip_runoff, aes(x = date, y = runoff_precip_ratio)) +
  geom_line(aes(color = month)) +
  theme_minimal() +
  scale_color_paletteer_d("rcartocolor::Bold") +
  scale_y_continuous(limits = c(0,1)) +
  scale_x_date(date_breaks = "5 year", date_labels = "%Y") +
  labs(x = "Year",
       y = "Runoff Ratio") +
  ggtitle("Santa Ynez Watershed Seasonal Runoff Ratio \n1983 - 2018")

seasonal_runoff_plot

# Takeaway: runoff ratio is highly variable in the watershed, even taking into consideration seasonality 

```

Seasonality graph for selected months (rainy season): 
```{r}

selected_seasons_plot <- monthly_precip_runoff %>% 
  filter(month == "01"|
           month == "02"|
           month == "11"|
           month == "12") %>% 
  ggplot(aes(x = date, y = runoff_precip_ratio)) +
  geom_line(aes(color = month)) +
  theme_classic() +
  scale_color_paletteer_d("rcartocolor::Bold") +
  scale_color_discrete("Month", labels = c("January", "February", "November", "December")) +
  scale_y_continuous(limits = c(0,1),
                     expand = c(0,0)) +
  scale_x_date(date_breaks = "5 year", date_labels = "%Y",
               expand = c(0,0)) +
  labs(x = "Year",
       y = "Runoff Ratio") +
  ggtitle("Santa Ynez Watershed Rainy Months Runoff Ratio (1983 - 2018)") +
  theme(plot.title = element_text(hjust = 0.5, size = 12))

selected_seasons_plot

```
